---
title: "Code Annex"
author: "ma206"
date: '2022-10-02'
output:
 rmdformats::readthedown:
 highlight: kate
 css: custom.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generic code for various statistical tests in R.

## Single Categorical Variable 

### Summary Statistic

```{r, eval = FALSE}
dataframe %>%
  count(variable) %>%
  mutate(prop = n/sum(n)) %>%
  adorn_totals()
```

Check whether or not validity conditions are met.

### Bargraph

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=variable)) +
  geom_bar()
```


### One-proportion z-test

Standard Deviation of the null distribution:

$sd_{null}=\sqrt{\frac{\pi_0(1-\pi_0)}{n}}$ 

```{r, eval = FALSE}
null =           # Enter the value of your Null Hypothesis Parameter
n =              # Enter the sample size
stat =           # Enter the value of your statistic
sd = sqrt(null*(1 - null)/n)  # Standard deviation of the null distribution
standardized_stat = (stat-null)/sd  # Standardized Statistic, z
```

Choose the $p$-value code that corresponds to the alternative hypothesis:

Less than hypothesis
```{r, eval = FALSE}
pvalue = pnorm(standardized_stat)
```

Greater than hypothesis
```{r, eval = FALSE}
pvalue = 1-pnorm(standardized_stat)
```

Two sided hypothesis
```{r, eval = FALSE}
pvalue = 2*(1-pnorm(abs(standardized_stat)))
```

### Confidence Interval

```{r, eval = FALSE}
siglevel =              #Enter your significance level (alpha)
multiplier = qnorm(1-siglevel/2) 
se = sqrt(stat*(1 - stat)/n)     # Standard Error
CI = c(stat-multiplier*se, stat+multiplier*se)  # Confidence Interval
```
$se=\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$ 

## Single Quantitative Variable

### Summary Statistic
```{r, eval = FALSE}
dataframe %>%
  summarise(mean = mean(variable),
            s = sd(variable),
            n = n())
```

### Histogram

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x = variable)) + 
  geom_histogram()
```

Check whether or not validity conditions are met.

### One-sample t-test

```{r, eval = FALSE}
null =           #Enter the value of your Null Hypothesis Parameter
n =              #Enter the sample size
stat =           #Enter the value of your statistic
s =              # Sample standard deviation (from summary statistics)
sd = s/sqrt(n)   # Standard deviation of the null distribution
standardized_stat = (stat-null)/sd # Standardized statistic, t
```

Standard Deviation of the null distribution:

$sd_{null}=s/\sqrt{n}$

Choose the $p$-value code that corresponds to the alternative hypothesis:

Less than hypothesis
```{r, eval = FALSE}
pvalue = pt(standardized_stat, n-1)
```

Greater than hypothesis
```{r, eval = FALSE}
pvalue = 1-pt(standardized_stat, n-1)
```

Two sided hypothesis

```{r, eval = FALSE}
pvalue = 2*(1-pt(abs(standardized_stat), n-1))
```

### Confidence Interval

```{r, eval = FALSE}
siglevel =              #Enter your significance level (alpha)
multiplier = qt(1-siglevel/2, n-1)
se = s/sqrt(n)          # Standard error
CI = c(stat-multiplier*se, stat+multiplier*se)  # Confidence Interval
```

Standard Error:

$se=s/\sqrt{n}$

## Comparing Two Proportions 
### Categorical Explanatory Variable, Categorical Response Variable

### Summary Statistic

```{r, eval = FALSE}
dataframe <- dataframe %>% 
  mutate(explanatoryvariable = as.factor(explanatoryvariable),
         responsevariable = as.factor(responsevariable)) # convert the categorical variables to a factors

dataframe %>%
  count(explanatoryvariable, 
        responsevariable) %>%
  pivot_wider(names_from = explanatoryvariable, 
              values_from = n) %>%
  adorn_totals(c("row", "col"))
```

Check whether or not validity conditions are met.

### Segmented Bar Graph

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x= explanatoryvariable,
             fill = responsevariable)) +
  geom_bar(position = position_fill())
```

### Two-proportion z-test

```{r, eval = FALSE}
null =           #Enter the value of your Null Hypothesis Parameter
n =              #Enter the sample size
stat =           #Enter the value of your statistic
successes_1 =     # number of successes in group 1
successes_2 =     # number of successes in group 2
n_1 =    # sample size of group 1     
n_2 =    # sample size of group 2
phat_1 = successes_1/n_1
phat_2 = successes_2/n_2
phat_t = (successes_1 + successes_2)/(n_1 + n_2)
stat = phat_1-phat_2 # ensure this matches your null hypothesis order
sd = sqrt(phat_t*(1-phat_t)*(1/n_1 + 1/n_2))
standardized_stat = (stat-null)/sd  # standardized statistic, z
```

Statistic: 

$\textrm{stat}=\hat{p_1}-\hat{p_2}$

Standard Deviation of the Null Distribution:

$sd_{null}=\sqrt{\hat{p}*(1-\hat{p})*(\frac{1}{n_1}+\frac{1}{n_2})}$

Choose the $p$-value code that corresponds to the alternative hypothesis:

Less than hypothesis
```{r, eval = FALSE}
pvalue = pnorm(standardized_stat)
```

Greater than hypothesis
```{r, eval = FALSE}
pvalue = 1-pnorm(standardized_stat)
```

Two sided hypothesis
```{r, eval = FALSE}
pvalue = 2*(1-pnorm(abs(standardized_stat)))
```

### Confidence Interval

```{r, eval = FALSE}
siglevel =              #Enter your significance level (alpha)
multiplier = qnorm(1-siglevel/2)
se = sqrt(phat_1*(1-phat_1)/n_1+phat_2*(1-phat_2)/n_2) # Standard Error
CI = c(stat-multiplier*se, stat+multiplier*se)  # Confidence Interval
```
$se=\sqrt{\frac{\hat{p_1}*(1 - \hat{p_1})}{n_1}+\frac{\hat{p_2}*(1 - \hat{p_2})}{n_2}}$

## Comparing Two Means
### Categorical Explanatory Variable, Quantitative Response Variable

### Summary Statistic
```{r, eval = FALSE}
dataframe <- dataframe %>% 
  mutate(catvariable = as.factor(catvariable)) # convert the categorical variable to a factor

dataframe %>%
  group_by(catvariable) %>%
  summarise(mean = mean(quantvariable),
            s = sd(quantvariable),
            n = n())
```

### Split Histogram

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=quantvariable)) + 
  geom_histogram() +
  facet_grid(catvariable~.)
```

Check whether or not validity conditions are met.

### Two-sample t-test

Statistic: 

$\textrm{stat}=\bar{x}_1-\bar{x}_2$

```{r, eval = FALSE}
null =           #Enter the value of your Null Hypothesis Parameter
n =              #Enter the sample size
xbar_1 =        # sample mean of group 1
xbar_2 =        # sample mean of group 2
s_1 =           # sample standard deviation of group 1
s_2 =           # sample standard deviation of group 2
n_1 =      # sample size of group 1
n_2 =        # sample size of group 2
stat = xbar_1 - xbar_2
sd = sqrt(s_1^2/n_1 + s_2^2/n_2)
standardized_stat = (stat-null)/sd # standardized statistic, t
```

Standard Deviation of the Null Distribution:

$sd_{null}=\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}$

Choose the $p$-value code that corresponds to the alternative hypothesis:

Less than hypothesis
```{r, eval = FALSE}
pvalue = pt(standardized_stat, n-2)
```

Greater than hypothesis
```{r, eval = FALSE}
pvalue = 1 - pt(standardized_stat, n-2)
```

Two sided hypothesis

```{r, eval = FALSE}
pvalue = 2*(1-pt(abs(standardized_stat), n-2))
```

### Confidence Interval

```{r, eval = FALSE}
siglevel =              #Enter your significance level (alpha)
multiplier = qt(1-siglevel/2, n-2)
se = sqrt(s_1^2/n_1 + s_2^2/n_2) # standard error
CI = c(stat-multiplier*se, stat + multiplier*se) # confidence interval
```
$se=\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}$

## Linear Regression

The model below is linear regression without any interaction terms.  Include as many variables as needed to be analyzed.
```{r, eval = FALSE}
lrmodel <- dataframe %>%
  lm(responsevar ~ var1 + var2, data = .)
summary(lrmodel)
```

The model below evaluates an interaction between the two variables.

```{r, eval = FALSE}
lrmodel <- dataframe %>%
  lm(responsevar ~ var1 * var2, data = .)
summary(lrmodel)
```

### Testing Validity Conditions

#### Validating **L**inearity and **E**qual Variance

```{r, eval = FALSE}
lrmodel%>%
  fortify(lrmodel$model)%>%
  ggplot(aes(x = .fitted, 
             y = .resid))+
  geom_point()+
  geom_hline(yintercept = 0)+
  labs( x = "Predicted Values", 
        y = "Residuals", 
        title = "Residuals vs. predicted values")
```

#### Validating the **I**ndependence condition

```{r, eval = FALSE}
lrmodel%>%
  fortify(lrmodel$model) %>%
  mutate(row = row_number()) %>%
  ggplot(aes(x = row, 
             y = .resid))+
  geom_point()+
  geom_hline(yintercept = 0)+
  labs(x = "Order of Occurence", 
       y = "Residuals", 
       title = "Residuals in Order of Occurence")
```

#### Validating the **N**ormality condition

```{r, eval = FALSE}
lrmodel%>%
  fortify(lrmodel$model)%>%
  ggplot(aes(x = .resid))+
  geom_histogram()+
  labs(x = "Residuals", 
       title = "Histogram of residuals")
```

## Visualizations

### Bargraph

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=as.factor(variable))) +
  geom_bar()
```

### Histogram

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=variable)) + 
  geom_histogram()
```

### Split Histogram
Note that the method below can be used to split any plot type up by a categorical variable's values.  This example is for a histogram.

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=quantvariable)) + 
  geom_histogram() +
  facet_grid(as.factor(catvariable)~.)
```

### Segmented Bar Graph

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=as.factor(explanatoryvariable), fill = as.factor(responsevariable))) +
  geom_bar(position = position_fill())
```


### Boxplot

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x = var1, y = var2)) +
  geom_boxplot()
```

### Scatterplot

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=var1, y=var2)) +
  geom_point()
```

### Scatterplot with Linear Regression Line

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x=var1, y=var2)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### Scatterplot with Linear Regression Interaction

```{r, eval = FALSE}
dataframe %>%
  ggplot(aes(x = var1, y = var2, color = var3)) +
  geom_point() +
  geom_smooth(method = "lm")
```

