---
title: "Three S approach with R"
subtitle: ""
author: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 10)
library(flipbookr)
```

---



class: inverse, left, bottom
background-image: url(https://images.unsplash.com/photo-1501641466388-c67e34ec767e?ixid=MXwxMjA3fDB8MHxzZWFyY2h8Mnx8b2xkJTIwZmFpdGhmdWx8ZW58MHx8MHw%3D&ixlib=rb-1.2.1&auto=format&fit=crop&w=500&q=60)
background-size: cover

# .Large[]

## .small[]
#### .tiny[Dr. Evangeline Reynolds | Meeting 4 | 2021-02-02 | Image credit: Emily Campbell, Upsplash]


---



## Observation

--

### We characterize the success of the dolphins in the experiment by calculating a statistic:

--

### proportion of successes

---

`r flipbookr::chunk_reveal("observed_statistic")`


```{r observed_statistic, include=F}
library(tidyverse)

c(1, 1, 1, 1, 1, 1, 1, 1,
  1, 0, 1, 1, 1, 1, 1, 1) ->
buzz_doris_experiment

buzz_doris_experiment %>% 
  mean() ->
buzz_doris_observed_prop
```


---

# But is this statistic consistent with the chance model?

--

## Let's look at some outcomes of individual trials under the chance model

---



`r flipbookr::chunk_reveal("individual_chance_trails", break_type = 10)`

```{r individual_chance_trails, include = FALSE}
0:1 %>% 
  sample(size = 16, 
         replace = T,
         prob = c(.5,.5))
```
---

## *Many* trials tell us about probability distribution in the chance model

--

### Writing a function will help us manage many trials


---

`r chunk_reveal("write_function")`

```{r write_function, include=F}
single_trial <- function(x){
  
  0:1 %>% 
  sample(size = 16, 
         replace = T,
         prob = c(.5,.5))
  
}

# function at work
single_trial() %>% 
  tibble(outcome = .) %>% 
  ggplot() + 
  aes(x = outcome) + 
  geom_bar()
```


---

`r chunk_reveal("many_trials")`

```{r many_trials, include = F}
set.seed(2022)
1:10000 %>% 
  tibble(trial = .) %>% 
  mutate(ho = 
           map(.x = trial, #input
               .f = single_trial #function
               )) %>% 
  mutate(num_successes = 
           map_dbl(ho, sum)) %>% 
  mutate(prop_success = 
           map_dbl(ho, mean)) ->
many_trials_chance_model  

many_trials_chance_model %>% 
  slice(1:12) %>% 
  unnest(cols = ho) %>% 
  ggplot() + 
  aes(x = ho) + 
  facet_wrap(~ trial) + 
  geom_bar() + 
  ggdirect::geom_text_count()
```

---

`r chunk_reveal("dist_statistic_trials")`



```{r dist_statistic_trials, include = F}  
ggplot(data = many_trials_chance_model) + 
  aes(x = prop_success) + 
  geom_rug(alpha = .2) + 
  geom_bar() + 
  ggxmean::geom_x_mean() + 
  ggxmean:::geom_x1sd(linetype = "dashed") + 
  ggxmean:::geom_x2sd(linetype = "dashed") + 
  ggxmean:::geom_x3sd(linetype = "dashed") + 
  ggxmean:::geom_normal_dist(height = 600, fill = "magenta") + 
  # situate observed in chance model trials
  geom_vline(xintercept = buzz_doris_observed_prop,
             color = "red") ->
plot
```

---

`r chunk_reveal("pvalue")`

```{r pvalue, include=F}
many_trials_chance_model %>% 
  mutate(as_or_more_extreme = prop_success >= 
           buzz_doris_observed_prop) %>% 
  pull(as_or_more_extreme) %>% 
  mean() # mean of true and false give proportion true


many_trials_chance_model %>%
  pull(prop_success) %>% 
  sd() ->
three_s_method_sd

# z score
(15/16 - .5)/.1249

```


<!-- adjust font size in this css code chunk, currently 80 -->

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 120%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```



