---
title: "One sample t test"
author: "ma206"
date: '2022-09-29'
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Step 0: Read question, and read in libraries and data

> On average, do people **overestimate** time spans?  A song snippet that  is *ten seconds* is played for respondents.  **On average** will respondents **overestimate** the length of the snippet played.  Answer this question given a 99% conficence level.

```{r cars, warning=F, message=F}
library(tidyverse)
library(janitor)

#read in data from internet
elapsedtime <- read_csv("https://raw.githubusercontent.com/EvaMaeRey/course_guide_datasets/main/chap2.ElapsedTime.csv")

# preview
head(elapsedtime)


# summarize
elapsedtime %>%   
  summarise(n = n(),
            xbar = mean(Time),
            s = sd(Time))
```

# Step 1: basic facts

## 1a Lay out null and alternative

$$ H_0: \mu = 10 $$
$$ H_A: \mu > 10 $$

## 1b record values

```{r}
mu = 10 # proposed parameter

n = 48 # number of observations
xbar = 13.71 # observed statistic (mean)
s = 6.50
```
## 1c plotting 

Plot the distribution of the outcomes

```{r, fig.height=2, fig.width=4, warning=F, message=F}
ggplot(data = elapsedtime) + 
  aes(x = Time) + 
  geom_dotplot()
```

# Step 2: Standarized Statistic, Z

## 2a Calculate

```{r}
t_numerator = xbar - mu
sd_null = s / sqrt(n)

t = t_numerator/sd_null

print(t)
```


## 2b Interpret

| value of t | strength |
|:--------------------:|---------------------|
|between −1.5 and 1.5 |little or no evidence against the null hypothesis| 
|below −1.5 or above 1.5|moderate evidence against the null hypothesis|
|below −2 or above 2|strong evidence against the null hypothesis|
|below −3 or above 3|very strong evidence against the null hypothesis|

# Step 3: P-Value

> The p-value is the probability of observing a result as or more extreme as the statistic (xbar), if the null is true. 

## 3a Calculate

Select the code according to the alternative hypothesis

|Alternative Hypothesis|R Code|
|:--------------------:|------|
|$$ > $$|`1-pt(t, df = n-1)`|
|$$ < $$|`pt(t, df = n-1)`|
|$$ \neq$$|`2*(1-pt(abs(t), df = n-1))`|

```{r}
pvalue = 1-pt(t, df = n-1)

print(pvalue)
```

## 3b Interpret


> Guidelines for evaluating strength of evidence from p-values

| Value | Interpretation|
|:---------------:|--------------------|
|0.10 < p-value|not much evidence against null hypothesis; null is plausible|
|0.05 < p-value ≤ 0.10|moderate evidence against the null hypothesis|
|0.01 < p-value ≤ 0.05|strong evidence against the null hypothesis|
|p-value ≤ 0.01|very strong evidence against the null hypothesis|

> The smaller the p-value, the stronger the evidence against the null hypothesis.

## 3c Compare to critical value

The critical value, alpha, is implied by the confidence level required.  The question requires 95% confidence.  This implies a .01 critical level (taking 1 minus .99).

We compare this to the p-value.  If the p-value is smaller, we can reject the null hypothesis.

```{r}
pvalue < .01 # if true reject the null
```

I will *reject* the null hypothesis. My p-value is 0.0001285488, which is less than alpha, the critical value. 

# Step 4: Confidence Interval

## 4a calculate

|**Confidence Interval**|**Multiplier (Categorical)**|Multiplier (Quantitative)|
|-------------------|-----------------------|-------------------------|
|90%|`qnorm(.95)`| `qt(.95,n-1)` or `qt(.95,n-2)`|
|95%|`qnorm(.975)`|`qt(.975,n-1)` or `qt(.975,n-2)`|
|99%|`qnorm(.995)`|`qt(.995,n-1)` or `qt(.995,n-2)`|


## 4b interpret

```{r}
se = s / sqrt(n)
m = qt(.975, df = n - 1)

lower_bound = xbar - m*se
upper_bound = xbar + m*se

print(lower_bound)
print(upper_bound)
```
